On the Use of K-Competitive Networks \\ for Writing Style Transfer

### 2017-12-14 dl_style_transfer/paper/main.tex

\author{
\alignauthor
Luca Leeser\\
       \affaddr{Cornell University}\\
       \email{ll698@cornell.edu}
\alignauthor
Yuji Akimoto\\
       \affaddr{Cornell University}\\
       \email{ya242@cornell.edu}
\and
\alignauthor 
Ryan Butler\\
       \affaddr{Cornell University}\\
       \email{rjb392@cornell.edu}
\alignauthor 
Cameron Ibrahim\\
       \affaddr{Cornell University}\\
       \email{cai29@cornell.edu}
}

@@

\begin{figure}[h]
\subfloat{\includegraphics[width = .3\linewidth]{gates-hall.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .3\linewidth]{starry_night.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .3\linewidth]{gates-starry.png}}
\caption{\textmd{Cornell's Gates Hall stylized in the form of Vincent Van Gogh's \textit{Starry Night}.}}
\end{figure}

@@

In this paper, we explore the use of $k$-competitive convolutional layers, a generalization of $k$-competitive layers designed for autoencoders in \cite{KATE}, which in turn are based on $k$-sparse layers as defined in \cite{MakhzaniF13}. These layers are integrated into a word-level convolutional neural network \cite{Kim14f}, and we perform a qualitative analysis of the quality of content and style representations generated by these networks. A key benefit of our method is that it does not require a training set of translations between writing styles, which are very difficult to obtain.

@@

% PRELIMINARY ANALYSIS
\section{Preliminary Analysis}
The majority of implementations of style transfer for images use the VGG-19 \cite{VGG19} architecture to generate content and style representations of an image. This network was trained to classify images into one of the one thousand classes of the ImageNet challenge dataset \cite{ILSVRC15} and therefore is able to generate informative feature representations of a very large variety of objects. Given the resources required to train a similarly robust classifier for text, we first explore the tradeoffs of using an underlying network that is trained to discriminate between a much smaller number of classes by training a VGG network with the CIFAR-10 dataset \cite{Krizhevsky09} (here, we use pre-trained weights provided by \cite{Liu15}). Although this modified network is far less generalizable in terms of classification, we found that the content representations provided by later layers of the network were still informative enough to reconstruct the original image, even for images containing objects not included in the 10 classes of CIFAR-10. By reconstruction, we refer to treating the input to the network as a variable, and minimizing $\ell_2$ loss between the input image and original image's feature representations in a certain layer.

\begin{figure}[h]
\subfloat{\includegraphics[width = .15\linewidth]{cornell-campus-small.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{anderson-small.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{boats.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{whale.jpg}}
\hspace{0.1cm} 
\subfloat{\includegraphics[width = .15\linewidth]{pizza.jpg}}
\hspace{0.1cm} \\
\subfloat{\includegraphics[width = .15\linewidth]{2600.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{anderson5800.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{boats2100.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{boats3200.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{pizza5000.png}}
\hspace{0.1cm} \\
\subfloat{\includegraphics[width = .15\linewidth]{corne9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{ander9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{boats9900.png}}
\hspace	{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{whale9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width = .15\linewidth]{pizza9900.png}}
\hspace{0.1cm}
\caption{\textmd{Several 32x32 images (top) reconstructed in the conv2\_2 layer (middle) and conv4\_2 layer (bottom) of VGG-19 trained on CIFAR-10 data; maximum 10000 iterations.}}
\end{figure}

@@

However, we also observe that this reduction in generalizability of the underlying network architecture is not without tradeoffs. When using the conv4\_2 layer at which to match the feature representations of two images (the layer used in Figure 1), the reconstructed images are mostly unrecognizable to the human eye, and takes a much larger number of iterations to produce something even remotely close to the original. While we do not dismiss the possibility that this loss in reconstruction accuracy could be mitigated by hyperparameter optimization, we conclude that high generalizability as a classifier is an important quality of the underlying architecture although not paramount.

% MODEL
\section{Model}
\subsection{Convolutional Neural Networks for Text Classification}
Although recurrent neural networks have proven adept at language recognition tasks, evaluating the overall content and style of written text is a task that is less dependent on sequential information, and a hidden state is likely too small to encode enough information for our purposes. We believe that a a convolutional neural network allows us to take advantage of the success of style transfer for images, in contrast to aforementioned approaches that treat style transfer as a machine translation problem.

@@

% EXPERIMENTS
\section{Experiments}
\subsection{Datasets}
We use the most prolific reviewers on Yelp\footnote{{\texttt{https://www.yelp.com/dataset/challenge}}} to generate a large amounts of text for various styles of writing. Although the reviews for a frequent user of Yelp may span a number of years, we make the assumption that a user's writing style remains constant over that period of time. This may not strictly be the case as language, particularly of the colloquial variety that is often found online, evolves over time and certain words serve as clear indicators of certain time periods. However, dealing with a time-varying vocabulary set is an extension that we have not yet considered. After identifying the most frequent reviewers, we tokenize all of their reviews using NLTK \cite{Loper02nltk:the} to generate a corpus.

\subsection{Comparison of Architectures}

\subsection{Hyperparameters and Training}

@@

% FURTHER DISCUSSION
\section{Future Extensions}

@@

% APPENDIX
%pagebreak

\begin{appendix}
%\href{https://github.com/CornellDataScience/Yelp-FA17/tree/master/dl_style_transfer}{Link} to Github repository for all code associated with this paper.

\end{appendix}
\end{document}

>>>>

\author{
\alignauthor
Cameron Ibrahim\\
       \affaddr{Cornell University}\\
       \email{cai29@cornell.edu}
\alignauthor
Yuji Akimoto\\
       \affaddr{Cornell University}\\
       \email{ya242@cornell.edu}
\and
\alignauthor 
Ryan Butler\\
       \affaddr{Cornell University}\\
       \email{rjb392@cornell.edu}
\alignauthor 
Luca Leeser\\
       \affaddr{Cornell University}\\
       \email{ll698@cornell.edu}
}

@@

\begin{figure}[h]
\subfloat{\includegraphics[width=.3\linewidth]{gates-hall.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.3\linewidth]{starry_night.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.3\linewidth]{gates-starry.png}}
\caption{\textmd{Cornell's Gates Hall stylized in the form of Vincent Van Gogh's \textit{Starry Night}.}}
\end{figure}

@@

In this paper, we explore the use of $k$-competitive convolutional layers, a generalization of $k$-competitive layers designed for autoencoders in \cite{KATE}, which in turn are based on $k$-sparse layers as defined in \cite{MakhzaniF13}. We hypothesize that this technique will allow us to mitigate the high-dimensional sparsity of textual data and reduce the problem of writing style transfer to the well-established field of image style transfer. We integrate $k$-competitive layers into preprocessing and training of a word-level convolutional neural network \cite{Kim14f}, and perform a qualitative analysis of the quality of content and style representations generated by these networks. A key benefit of our method is that it does not require a training set of translations between writing styles, which are very difficult to obtain.

@@

% PRELIMINARY ANALYSIS
\section{Preliminary Analysis}
The majority of implementations of style transfer for images use the VGG-19 \cite{VGG19} architecture to generate content and style representations of an image. This network was trained to classify images into one of the one thousand classes of the ImageNet challenge dataset \cite{ILSVRC15} and therefore is able to generate informative feature representations of a very large variety of objects. Given the resources required to train a similarly robust classifier for text, we first explore the tradeoffs of using an underlying network that is trained to discriminate between a much smaller number of classes by training a VGG network with the CIFAR-10 dataset \cite{Krizhevsky09} (here, we use pre-trained weights provided by \cite{Liu15}). Although this modified network is far less generalizable in terms of classification, we found that the content representations provided by later layers of the network were still informative enough to reconstruct the original image, even for images containing objects not included in the ten classes of CIFAR-10. By reconstruction, we refer to treating the input to the network as a variable, and minimizing $\ell_2$ loss between the input image and original image's feature representations in a certain layer.

\begin{figure}[h]
\subfloat{\includegraphics[width=.15\linewidth]{cornell-campus-small.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{anderson-small.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{boats.jpg}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{whale.jpg}}
\hspace{0.1cm} 
\subfloat{\includegraphics[width=.15\linewidth]{pizza.jpg}}
\hspace{0.1cm} \\
\subfloat{\includegraphics[width=.15\linewidth]{2600.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{anderson5800.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{boats2100.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{boats3200.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{pizza5000.png}}
\hspace{0.1cm} \\
\subfloat{\includegraphics[width=.15\linewidth]{corne9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{ander9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{boats9900.png}}
\hspace	{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{whale9900.png}}
\hspace{0.1cm}
\subfloat{\includegraphics[width=.15\linewidth]{pizza9900.png}}
\hspace{0.1cm}
\caption{\textmd{Several 32x32 images (top) reconstructed in the conv2\_2 layer (middle) and conv4\_2 layer (bottom) of VGG-19 trained on CIFAR-10 data; maximum 10000 iterations.}}
\end{figure}

@@

However, we also observe that this reduction in generalizability of the underlying network architecture is not without tradeoffs. When using conv4\_2 as the layer at which to match the feature representations of two images (the layer used in Figure 1), the reconstructed images are mostly unrecognizable, and it takes a much larger number of iterations to produce something even remotely close to the original. However, we believe that there is enough signal even within the unrecognizable images, and with further hyperparameter optimization we will be able to mitigate this loss in reconstruction accuracy. This justifies the use of networks for author classification that have been trained on a limited number of classes as a starting point.

% MODEL
\section{Model}
\subsection{Convolutional Neural Networks for Text Classification}
Although recurrent neural networks have proven adept at language recognition tasks, evaluating the overall content and style of written text is a task that is less dependent on sequential information (outside the scope of a receptive field), and a hidden state is likely too small to encode enough information for our purposes. We believe that a a convolutional neural network allows us to take advantage of the success of style transfer for images, in contrast to aforementioned approaches that treat style transfer as a machine translation problem.

@@

% EXPERIMENTS
\section{Experiments}
\subsection{Datasets}
We use the most prolific reviewers on Yelp\footnote{{\texttt{https://www.yelp.com/dataset/challenge}}} to generate a large amounts of text for various styles of writing. Although the reviews for a frequent user of Yelp may span a number of years, we make the assumption that a user's writing style remains constant over that period of time. This may not strictly be the case as language, particularly of the colloquial variety that is often found online, evolves over time and certain words serve as clear indicators of certain time periods. However, dealing with a time-varying vocabulary set is an extension that we have not yet considered. After identifying the most frequent reviewers, we tokenize all of their reviews using NLTK \cite{Loper02nltk:the} to generate a corpus. To create a baseline task, we also create a corpus of Shakespearean works from Project Gutenberg\footnote{{\texttt{https://www.gutenberg.org/}}}.

\subsection{Comparison of Architectures}
We explore the use of $k$-competitive layers in two settings: in an autoencoder to generate word embeddings, and as a non-linearity after convolutional layers.

\subsubsection{K-Competitive Autoencoder for Embeddings}
In natural language processing tasks it is common encode textual data into a sparse space before generating embeddings that project these word or sentence-level vectors into a lower dimensional space. The $k$-competitive autoencoder for text (KATE) proposed in \cite{KATE} generates word embeddings that compare favorably to the more commonly used word2vec \cite{Mikolov13}, and we explore its use as a preprocessing step for an convolutional author classifier network. The two variants that we explore are in the initial sparse representation of training data: we compare between the use of a sentence-level log-normalized word count vectors similar to \cite{KATE} and word-level one-hot encoded vectors.

Since the autoencoder is used to generate vector embeddings of words, it can be argued that using sentence-level embeddings as training data creates a model that is asked to perform different tasks at training and test time. At the same time, these log-normalized word count vectors contain entries between 0 and 1, and can therefore be interpreted as probabilities of potentially overlapping classes. Compared to the one-hot encoding approach that treats words as belonging in mutually exclusive classes, this approach more accurately reflects the reality that words often have overlapping meanings. The amount of training data produced from a single passage of text depends on the relative magnitudes of the number of unique words and number of unique sentences, but given an even ratio the word count approach likely produces more useful data - a single word is much more likely to be meaningless noise than a sentence.

Consistent with the probabilistic interpretation mentioned above, we swap the sigmoid activation function used in the word count approach for a softmax function when using one-hot encoding. We explore the tradeoffs between these two approaches, and quantitatively assess their performance by comparing reconstruction accuracy of the autoencoder, that is, the $\ell_2$ loss between the input vector and the decoded vector of its low dimensional representation.

\subsubsection{K-Competitive Layer as Activation Function}
Both convolutional and fully connected $k$-competitive layers are non-linear functions and are therefore appropriate as activation functions. Since they set the activations of the least active neurons to 0, these act similarly to a rectified linear unit with an adaptive threshold. As with ReLU, this property enforces sparsity among the coefficients and therefore acts both as a regularizer and non-linearity. Meanwhile, the allocation of energy from ``loser'' to ``winner'' nodes prevents these neurons from becoming entirely dead and preserves gradient flow in backpropagation. Since the baseline convolutional architecture is shallow and only has one convolutional layer (albeit many filters), we add one $k$-competitive layer following the convolutional layer, with a ReLU layer following it to preserve the non-linearity at inference time (the $k$-competitive layer is defined as the identity function at test time in \cite{KATE}). The classification tasks are: the baseline tasks of differentiating Shakespearean text from the most prolific Yelp reviewer; differentiating between the two most prolific Yelp reviewers.

We compare the performance of the following model configurations on two author classification tasks: the baseline convolutional network with one-hot encoded vectors as input; the baseline convolutional neural network with KATE embedded vectors as input; a convolutional neural network with a $k$-competitive layer taking one-hot encoded vectors as input.

\subsection{Hyperparameters and Training}
All models were implemented in TensorFlow and trained on a TITAN Xp GPU. Our implementation of the convolutional neural network from \cite{Kim14f} is credited to Denny Britz\footnote{{\texttt{https://github.com/dennybritz}}}. Both variations of the $k$-competitive autoencoder were trained with a bottleneck layer of size 128, $k=32$, amplification constant $\alpha=6.26$ \cite{KATE}, learning rate 0.001, 100 epochs, and a batch size of 128 using the Adam optimizer \cite{KingmaB14}. All variations of the convolutional network were trained with dropout rate 0.5, $\ell_2$ regularization parameter $\lambda=0$, learning rate 0.001, 200 epochs, and a batch size of 64 using the Adam optimizer. Variations with the $k$-competitive layer used $k=32, 64$ and $\alpha=6.26$.

@@

% FURTHER DISCUSSION
\section{Future Extensions}
Varying length of what is considered a single ``document'' on which to generate word count vector for KATE.

@@

% APPENDIX
\begin{appendix}
All code associated with this paper can be found here.
\end{appendix}

\balance

\end{document}
